---
title: "Predicting Solar-Readiness of NYC Municipal Buildings"
author: "Qingquan Li"
date: "05/08/2026"
output:
  beamer_presentation:
    theme: "Madrid"
    colortheme: "seahorse"
    fonttheme: "structurebold"
    slide_level: 2
    keep_tex: false
header-includes:
  - \usepackage{booktabs}
  - \usepackage{graphicx}
  - \setbeamertemplate{navigation symbols}{}
  - \setbeamertemplate{footline}[frame number]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 7, fig.height = 3.8, fig.align = "center",
                      out.width = "95%")
library(tidyverse)
library(scales)
library(knitr)
```

```{r data-prep}
# data source: https://data.cityofnewyork.us/City-Government/City-of-New-York-Municipal-Solar-Readiness-Assessm/cfz5-6fvh/about_data
solar_data <- "https://raw.githubusercontent.com/qingquan-li/data606-project/refs/heads/main/data/City_of_New_York_Municipal_Solar-Readiness_Assessment_(Local_Law_24_of_2016)_data_20260413.csv"
raw <- read_csv(solar_data, show_col_types = FALSE)

# Focus on assessed buildings with Solar-Readiness = Yes or No
solar <- raw %>%
  filter(`Solar-Readiness Assessment` %in% c("Yes", "No"),
         Borough != "Upstate") %>%
  mutate(
    solar_ready = factor(`Solar-Readiness Assessment`, levels = c("No", "Yes")),
    solar_num = ifelse(solar_ready == "Yes", 1, 0),
    sqft = as.numeric(gsub("[^0-9.]", "", `Total Gross Square Footage`)),
    log_sqft = log(sqft),
    roof_condition = factor(`Roof Condition`,
      levels = c("Poor", "Between Fair and Poor", "Fair",
                 "Between Good and Fair", "Good",
                 "Unknown", "Under Construction")),
    roof_age = `Roof Age`,
    borough = factor(Borough)
  ) %>%
  filter(!is.na(sqft))

# Subset for modeling: only known roof conditions
model_data <- solar %>%
  filter(roof_condition %in% c("Poor","Between Fair and Poor","Fair",
                                "Between Good and Fair","Good"))
```

## Abstract

\small
New York City mandates biennial solar-readiness assessments of municipal buildings under Local Law 24 of 2016. This study examines whether observable building characteristics -- roof condition, roof age, building size (gross square footage), and borough -- are associated with a building's solar-readiness determination.

Using `r nrow(model_data)` assessed NYC municipal buildings from 2020 and 2022 reporting cycles, we fit a logistic regression model with solar-readiness (Yes/No) as the binary response. Results show that roof condition and building size are the strongest predictors. Buildings with "Good" roofs have 4.2x the odds of being solar-ready versus "Poor" roofs (p < 0.001). Each unit increase in log(sqft) doubles the odds (OR = 2.02, p < 0.001). The Bronx has 3x the odds of Manhattan (p < 0.001).

A chi-squared test confirms a significant association between roof condition and solar-readiness ($\chi^2$ = 68.26, df = 4, p < 0.001). Targeted roof maintenance could meaningfully expand the city's rooftop solar capacity.

## Overview: Data Context

- **Source:** NYC Open Data (DCAS), Local Law 24 of 2016
- **Collection:** Biennial census of city-owned buildings > 10,000 sqft
- **Study type:** Observational (no experimental manipulation)
- **Cases:** `r nrow(model_data)` assessed buildings (2020 & 2022 cycles)

\vspace{0.3cm}

**Research Question:** Is there a statistically significant association between a NYC municipal building's roof condition, building size, and borough -- and whether it is assessed as solar-ready?

## Overview: Variables

**Dependent (Response):**

- `solar_ready` -- Solar-Readiness Assessment (Yes/No) -- *categorical, binary*

\vspace{0.3cm}

**Independent (Explanatory):**

- `roof_condition` -- Roof Condition (Poor to Good) -- *ordinal categorical*
- `roof_age` -- Roof Age category -- *ordinal categorical*
- `sqft` -- Total Gross Square Footage -- *quantitative, continuous*
- `borough` -- NYC Borough -- *nominal categorical*

## Summary Statistics

```{r summary-stats}
summary_tbl <- model_data %>%
  group_by(roof_condition) %>%
  summarise(
    N = n(),
    Solar_Ready = sum(solar_num),
    Rate = paste0(round(mean(solar_num) * 100, 1), "%"),
    .groups = "drop"
  )
kable(summary_tbl, col.names = c("Roof Condition", "N", "Solar-Ready", "Rate"),
      align = c("l", "r", "r", "r"), booktabs = TRUE)
```

\vspace{0.2cm}
\small
Overall: **`r nrow(model_data)`** buildings | **`r sum(model_data$solar_num)`** solar-ready (**`r round(mean(model_data$solar_num)*100, 1)`%**) | Mean sqft: **`r comma(round(mean(model_data$sqft)))`** | Median sqft: **`r comma(round(median(model_data$sqft)))`**

## Solar-Readiness by Roof Condition

```{r chart-roof-condition, fig.height=4}
roof_rates <- model_data %>%
  group_by(roof_condition) %>%
  summarise(rate = mean(solar_num) * 100, .groups = "drop") %>%
  mutate(short = c("Poor", "Fair-Poor", "Fair", "Good-Fair", "Good"))

ggplot(roof_rates, aes(x = fct_inorder(short), y = rate, fill = short)) +
  geom_col(width = 0.7, show.legend = FALSE) +
  geom_text(aes(label = paste0(round(rate, 1), "%")),
            vjust = -0.5, fontface = "bold", size = 4) +
  scale_fill_manual(values = c("#E74C3C","#E67E22","#F1C40F","#82C91E","#27AE60")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Roof Condition", y = "% Solar-Ready",
       title = "Solar-Readiness Rate by Roof Condition") +
  theme_minimal(base_size = 13) +
  theme(panel.grid.major.x = element_blank())
```

## Building Size by Solar-Readiness

```{r chart-sqft, fig.height=4}
ggplot(model_data, aes(x = solar_ready, y = sqft, fill = solar_ready)) +
  geom_boxplot(width = 0.5, show.legend = FALSE) +
  scale_fill_manual(values = c("No" = "#BDC3C7", "Yes" = "#2980B9")) +
  scale_y_log10(labels = comma) +
  labs(x = "Solar-Ready?", y = "Gross Square Footage (log scale)",
       title = "Building Size by Solar-Readiness Status") +
  theme_minimal(base_size = 13)
```

## Solar-Readiness by Borough

```{r chart-borough, fig.height=4}
boro_rates <- model_data %>%
  group_by(borough) %>%
  summarise(n = n(), rate = mean(solar_num) * 100, .groups = "drop")

ggplot(boro_rates, aes(x = borough, y = rate, fill = borough)) +
  geom_col(width = 0.7, show.legend = FALSE) +
  geom_text(aes(label = paste0(round(rate, 1), "%\n(n=", n, ")")),
            vjust = -0.3, size = 3.5) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(x = "Borough", y = "% Solar-Ready",
       title = "Solar-Readiness Rate by Borough") +
  theme_minimal(base_size = 13) +
  theme(panel.grid.major.x = element_blank())
```

## Solar-Readiness by Roof Age

```{r chart-roof-age, fig.height=4}
age_rates <- solar %>%
  filter(!roof_age %in% c("Unknown", "Under Construction")) %>%
  mutate(roof_age = case_when(
    roof_age == "< 2 years"      ~ "< 2 yrs",
    roof_age %in% c("2 to 4 years", "2 to 4 Years") ~ "2-4 yrs",
    roof_age == "5 to 9 years"   ~ "5-9 yrs",
    roof_age == "10 to 14 years" ~ "10-14 yrs",
    roof_age == "15 to 19 years" ~ "15-19 yrs",
    roof_age == "> 20 years"     ~ "> 20 yrs"
  )) %>%
  mutate(roof_age = factor(roof_age,
    levels = c("< 2 yrs","2-4 yrs","5-9 yrs","10-14 yrs","15-19 yrs","> 20 yrs"))) %>%
  group_by(roof_age) %>%
  summarise(rate = mean(solar_num) * 100, .groups = "drop")

ggplot(age_rates, aes(x = roof_age, y = rate)) +
  geom_col(fill = "#2980B9", width = 0.7) +
  geom_text(aes(label = paste0(round(rate, 1), "%")),
            vjust = -0.5, fontface = "bold", size = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Roof Age", y = "% Solar-Ready",
       title = "Solar-Readiness Rate by Roof Age") +
  theme_minimal(base_size = 13) +
  theme(panel.grid.major.x = element_blank())
```

## Chi-Squared Test: Roof Condition vs Solar-Readiness

**Hypotheses:**

- $H_0$: Roof condition and solar-readiness are independent (no association).
- $H_a$: There is a significant association between roof condition and solar-readiness.

```{r chi-squared}
model_data_clean <- droplevels(model_data)
ct <- table(model_data_clean$roof_condition, model_data_clean$solar_ready)
chi_test <- chisq.test(ct)
chi_test
```

\vspace{0.3cm}
**Conclusion:** With p < 0.001, we **reject $H_0$**. There is strong evidence of a statistically significant association between roof condition and solar-readiness at the $\alpha = 0.05$ level.

## Logistic Regression Results

\small
Model: `solar_ready ~ log(sqft) + roof_condition + borough` (Ref: Poor roof, Manhattan)

```{r logistic-regression}
model_data <- droplevels(model_data)
model_data$roof_condition <- relevel(model_data$roof_condition, ref = "Poor")
model_data$borough <- relevel(model_data$borough, ref = "Manhattan")

fit <- glm(solar_num ~ log_sqft + roof_condition + borough,
           data = model_data, family = binomial)

coef_tbl <- summary(fit)$coefficients
or_tbl <- data.frame(
  Predictor = c("(Intercept)", "log(sqft)",
                "Roof: Fair-Poor", "Roof: Fair", "Roof: Good-Fair", "Roof: Good",
                "Boro: Bronx", "Boro: Brooklyn", "Boro: Queens", "Boro: Staten Isl."),
  Coef = round(coef_tbl[, 1], 3),
  OR = round(exp(coef_tbl[, 1]), 3),
  z = round(coef_tbl[, 3], 3),
  p = ifelse(coef_tbl[, 4] < 0.001, "<0.001", round(coef_tbl[, 4], 4)),
  row.names = NULL
)
kable(or_tbl, col.names = c("Predictor", "Coef", "Odds Ratio", "z", "p-value"),
      booktabs = TRUE, align = c("l","r","r","r","r"))
```

## Logistic Regression: Model Fit

```{r model-fit}
pseudo_r2 <- round(1 - fit$deviance / fit$null.deviance, 4)
aic_val <- round(AIC(fit), 1)
```

- **McFadden's Pseudo $R^2$:** `r pseudo_r2` -- the model explains about 13% of the variation in solar-readiness, a modest but meaningful fit using only physical building attributes.
- **AIC:** `r aic_val`
- **LLR p-value:** < 0.001 (the overall model is statistically significant)
- **n:** `r comma(nrow(model_data))`

\vspace{0.3cm}

**Interpretation of Pseudo $R^2$:** Unlike OLS $R^2$, McFadden's Pseudo $R^2$ values between 0.2--0.4 are considered excellent. Our value of 0.13 suggests that unmeasured factors (structural capacity, electrical infrastructure, shading) also influence solar-readiness, but the physical building characteristics in our model still provide useful predictive power.

## Key Findings

1. **Roof Condition matters most:** Buildings with "Good" roofs have **4.2x** the odds of being solar-ready vs. "Poor" roofs (p < 0.001).

2. **Larger buildings are more viable:** Each unit increase in log(sqft) **doubles** the odds of solar-readiness (OR = 2.02, p < 0.001).

3. **Borough effect:** The Bronx has **3x** the odds of Manhattan (p < 0.001), possibly reflecting newer roof stock or building types.

4. **Model fit:** McFadden's Pseudo $R^2$ = 0.13 indicates a modest but meaningful fit -- reasonable given we are using only physical building attributes.

## Conclusion

**Why This Matters:**
NYC has committed to carbon neutrality by 2050. Understanding which building characteristics predict solar-readiness helps the city prioritize roof maintenance and target capital investments to expand rooftop solar across its municipal portfolio.

\vspace{0.3cm}

**Limitations:**

- Only 3.6% of buildings are solar-ready -- class imbalance may affect model sensitivity
- Observational design -- cannot establish causal relationships
- Pseudo $R^2$ = 0.13 suggests unmeasured factors (structural, electrical, shading) also drive solar-readiness
- Only municipal buildings -- results may not generalize to private buildings

\vspace{0.3cm}
\tiny
Data: NYC Open Data -- City of New York Municipal Solar-Readiness Assessment (Local Law 24 of 2016) | \url{https://data.cityofnewyork.us/City-Government/City-of-New-York-Municipal-Solar-Readiness-Assessm/cfz5-6fvh}
